<!--
Update the file paths as needed throughout
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Import**
***

Now that you have set up this project with {{< glossary "Git">}} and {{< glossary "GitHub">}}, you can start working on the study comparison in R. Open RStudio or Positron or your preferred R IDE to the "infant_gut_analysis" folder. We will start by taking a look at the data. Take a look at the {{< glossary "README">}} file to determine which file contains commands that can read in the raw data. Open this file and run the code. 

<details> <summary> Click here to see the code included in this file. </summary>

::: {.rcode}
```r
library(readxl) # for reading data from supplementary table
library(dplyr) # for combining data across studies 

# read tabs for each study from Table S1 from Wang et al. supplementary 
# information, saved as "meta_analysis_metadata.xlsx"
study_data <- lapply(3:10, function(i) {
  read_xlsx("raw_data/Table S1.xlsx", sheet = i)
}) 

# set all variables as characters in order to combine data across studies
# original data had some variables in different data types across studies
study_data_clean <- lapply(study_data, function(df) {
  df %>% mutate(across(everything(), as.character))
})

# combine lists for each study into a single data frame
sample_metadata <- bind_rows(study_data_clean)

# save data frame as a .rda file
save(sample_metadata, file = "processed_data/sample_metadata.rda")
```
:::

</details>

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
How many observations are included in our data set? How many variables are included in our data set? 
<details> <summary> Hint </summary>

Data is imported in the file "01_pull_study_data.R". You can find out how many rows are in a data frame with the function `nrow(name_of_data_frame)`. You can find out how many columns are in a data frame with the function `ncol(name_of_data_frame)`. 

</details>
<details><summary>Answer</summary>

```{r, echo = FALSE}
load(here::here("data", "imported", "sample_metadata.rda"))
```

::: {.rcode}
```{r}
nrow(sample_metadata)
ncol(sample_metadata)
```
:::

There are 1,499 observations and 20 variables in the data frame.
</details>
::::
:::

***
