<!--
Update the file paths as needed throughout
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Visualization**
***

## Create a new branch for the analysis

Note: If we describe branches in the "Introduction to Git and GitHub" section then we don't need to here. 

## Visualize data

to-do: break this into smaller pieces, provide more information, and give learner space to figure things out on their own, with hints and solutions in drop-down boxes. See original case studies for examples here! 

to-do: have the learner open a .R or .Rmd file for this code.
Question: preference on .R and .Rmd files for this?

Load data 

::: {.rcode}
```{r}
load(here::here("data", "wrangled", "clean_sample_metadata.rda"))
```
:::

Count samples by study and category (mother vs infant)

::: {.rcode}
```{r}
sample_counts <- clean_metadata %>%
  group_by(Study, Category) %>%
  count(name = "total")
sample_counts
```
:::

Make a bar plot of this data 

::: {.rcode}
```{r}
sample_counts %>%
  ggplot(aes(x = Study, y = total, fill = Category)) +
  geom_col(position = position_dodge())
```
:::

Order the bar plot by increasing sample size and add theme and labels

::: {.rcode}
```{r}
study_order <- sample_counts %>%
  group_by(Study) %>%
  summarise(total_both = sum(total)) %>%
  arrange(total_both) %>%
  pull(Study)
sample_counts %>%
  mutate(Study = factor(Study, levels = study_order)) %>%
  ggplot(aes(x = Study, y = total, fill = Category)) +
  geom_col(position = position_dodge()) +
  theme_minimal() + 
  labs(y = "Number of samples",
       x = "Study",
       title = "Number of Samples by Study and Category") + 
  theme(plot.title = element_text(hjust = 0.5))
```
:::

Save plot in figures folder. 

::: {.rcode}
```{r, eval = FALSE}
ggsave("figures/study_sizes_by_category.png", height = 6, width = 9)
```
:::

Now look into time 

::: {.rcode}
```{r}
clean_metadata %>%
  filter(!sampling_before_birth) %>%
  ggplot(aes(x = sampling_day, fill = Category, group = Category)) +
  geom_histogram(position = position_dodge(), bins = 20) +
  facet_wrap(~Study, scale = "free_y") 
```
:::

Order facets the same as in bar plot, add theme and labels 

::: {.rcode}
```{r}
clean_metadata %>%
  mutate(Study = factor(Study, levels = study_order)) %>%
  filter(!sampling_before_birth) %>% # remove samples taken from mothers while pregnant 
  ggplot(aes(x = sampling_day, fill = Category, group = Category)) +
  geom_histogram(position = position_dodge(), bins = 20) +
  facet_wrap(~Study, scale = "free_y") +
  labs(x = "Days since birth at sampling",
       y = "Count",
       title = "Distribution of sampling day across studies") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
:::

Save plot in figures folder

::: {.rcode}
```{r, eval = FALSE}
ggsave("figures/sampling_timepoints.png", height = 6, width = 9)
```
:::

Now that you have two plots that compare the studies, this is a good time to commit your current work. 

Since you've performed the requested analysis, make sure to document any new files in this repository in the README file. This is another good time to commit any changes. 

***
