<!--
Update the file paths as needed throughout
-->

<!--
Some case studies may wish to merge Data Exploration with this Wrangling subsection
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Wrangling**
***

Data wrangling includes cleaning data to fix mistakes and reorganizing data to make it easier to use in visualizations and analyses. Use the {{< glossary "README" >}} to identify which file performs data wrangling. Open this file and run the code. 

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
What are the earliest and latest days that samples were taken on in the data set?  
<details> <summary> Hint </summary>

Data are wrangled in the file "02_wrangle_data.R". You can use the functions `max()` and `min()` to get the maximum and minimum values in a vector. The variable in the cleaned data set that gives the sampling day is called `sampling_day`. 

</details>
<details><summary>Answer</summary>

```{r, echo = FALSE}
load(here::here("data", "wrangled", "clean_sample_metadata.rda"))
```

::: {.rcode}
```{r}
min(clean_metadata$sampling_day, na.rm = TRUE)
max(clean_metadata$sampling_day, na.rm = TRUE)
```

The earliest day that a participant was sampled after birth was $0$, and the latest day was $480$. 

:::


</details>
::::
:::

***
