<!--
Update the file paths as needed throughout
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Import**
***

Now that you have set up this project with {{< glossary "Git" >}} and {{< glossary "GitHub" >}}, you can start working on the study comparison in R. Open RStudio or Positron or your preferred R IDE to the "infant_gut_analysis" folder. We will start by taking a look at the data. To do this, use the README.md file to determine which file reads in the raw data. Open this file and run the code. 

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
How many observations are included in our data set? How many variables are included in our data set? 
<details> <summary> Hint </summary>

Data is imported in the file "01_pull_study_data.R". You can find out how many rows are in a data frame with the function `nrow(name_of_data_frame)`. You can find out how many columns are in a data frame with the function `ncol(name_of_data_frame)`. 

</details>
<details><summary>Answer</summary>

```{r, echo = FALSE}
load(here::here("data", "imported", "sample_metadata.rda"))
```

::: {.rcode}
```{r}
nrow(sample_metadata)
ncol(sample_metadata)
```
:::

There are $1499$ observations and $20$ variables in the data frame.
</details>
::::
:::

***
