
# **Data Visualization**
***

We will now generate visualizations to compare the {{< glossary "meta-analysis">}} study cohorts.

First, we will make sure that we are on the correct branch. We want to use a development branch for this analysis, which we will request to merge back into the main branch once the analysis is ready. We already created the branch "study_comparison" in the [introduction to Git and GitHub section](https://www.opencasestudies.org/ocs-bio-version-control/#branches), so we just need to make sure that we are working on that branch.

::: {.panel-tabset}

#### GitHub Desktop 

- In the upper panel, under "Current Branch", look to see if it says the current branch is "study_comparison" or "main" or something else.
  - If it says "study_comparison", then you are already working on the correct branch.
  - If it says "main", then you can click the drop-down button next to "Current Branch" and select "study_comparison".
    - If "study_comparison" does not appear, you can click the "New Branch" button to create it. 

```{r, fig.alt="Creating first repository", out.width = "100%", echo = FALSE, fig.align = "center"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f94a4bdf40_0_170#slide=id.g3f94a4bdf40_0_170")
```

#### Command line 

Open the command line and navigate to your repository. Use the `git branch` command. 

:::: {.shell}
```bash
git branch
```
::::

This should list the branches for your repository, with the current branch highlighted and marked with an asterisk (*). If the highlighted branch is "study_comparison", then you are already working on the correct branch. If the highlighted branch is "main" or something else, you can switch over to "study_comparison" using the `git checkout` command. 

If "study_comparison" appears in your list of branches, you can move to it with the following command:

:::: {.shell}
```bash
git checkout study_comparison
```
::::

However, if "study_comparison" doesn't appear in your list of branches, you can create it with `git checkout -b`. 

:::: {.shell}
```bash
git checkout -b study_comparison
```
::::
:::

Because this is a new component of the analysis, we will create a new file in the "infant_gut_analysis" folder for this code. 

:::{.reproducibility_box}
::::{.reproducibility_box_header}
Reproducibility considerations
::::
::::{.inner_block}
There are different types of files that you can create to run R code. The most basic file is a .R script, which is used for the scripts that we ran in the data import and wrangling sections: "01_pull_study_data.R" and "02_wrangle_data.R". An R script assumes that all contents are R code, except for comments (which are preceded by `#`). 

Another file type is a notebook-style document, which combines text components with code components. The code is broken into pieces, which are referred to as "code chunks". Two common types of these files are R Markdown files (.Rmd) and quarto files (.qmd). These files can be rendered, which means that all code chunks are run in order (typically in a fresh R session), and any output (calculations, tables, plots) is included in the resulting document. These notebook-style formats also make it easier to explain what each step of an analysis is doing, improving both readability and reproducibility by clearly linking code, output, and interpretation in a single document.
::::
:::

To start, we will create a new quarto document and call it "03_visualize_study_cohorts.qmd". Make sure to give your quarto document an appropriate title. 

In this section, we will go through the steps to make two visualizations to compare the study cohorts. As you go, include the relevant code (as well as explanations of what the code is doing) into your quarto document. When you are done, you or someone else with access to the data and your document, should be able to run the code and create the visualizations. 

In your quarto document, you should start by making a code chunk to load the required packages. 

<details>
<summary>Adding a code chunk to a quarto file</summary>

::: {.panel-tabset}

#### Manually

To add a new R code chunk, go to a new line and enter: ```{r}

Then, type your code, starting on the following line. To end a code chunk, go to another new line and enter: ```

#### In RStudio

You can also add a code chunk in RStudio by navigating to "Code" -> "Insert Chunk". There is also a keyboard shortcut: `Ctrl + Alt + I` in Windows and Linux, and `Cmd + Option + I` on Mac. 

#### In Positron

In Positron there is a keyboard shortcut to add a code chunk: `Ctrl + Shift + I` in Windows and Linux, and `Cmd + Shift + I` on Mac. 

:::
</details>

**Include this code in your .qmd file**, to load the packages that we'll need. 

::: {.rcode}
```{r, message = FALSE, warning = FALSE}
library(here)
library(dplyr)
library(ggplot2)
```
:::

Next, we will load the wrangled data. We will use the function `here()`, which creates relative file paths. 

<details>
<summary>More information about the `here()` function:</summary>
The `here()` function creates file paths that always start from the main project folder (the top-level folder that contains all of your project files). In a Git-tracked repository, this is the folder that contains the `.git` directory, which was created when you initialized the repository. Using relative file paths (i.e., paths that only reference folders within your project) makes your code more portable and easier for collaborators to run. In contrast, absolute file paths depend on the specific file structure of your personal computer and will not necessarily work on other machines.
</details>

We want to load data from the file "clean_sample_metadata.rda", contained within the "processed_data" folder. **Include this code in your .qmd file** to load the data.

```{r, echo = FALSE}
load(here::here("data", "wrangled", "clean_sample_metadata.rda"))
```

::: {.rcode}
```{r, eval = FALSE}
load(here::here("processed_data", "clean_sample_metadata.rda"))
```
:::

Now that you have created a new file and loaded packages and data, this is a good time to track this new file and {{< glossary "commit">}} your work. 

::: {.panel-tabset}

#### GitHub Desktop 

Navigate back to {{< glossary "GitHub Desktop">}}. You should see your new file name "03_visualize_study_cohorts.qmd" with a green plus next to it, meaning that this is a new file. The check in the checkbox means that this file is staged. Add an appropriate commit message and commit this file. 

#### Command line 

Open the {{< glossary "command line">}} and navigate to your repository. Recall that you can stage the new file (tell Git to track it with):

:::: {.shell}
```bash
git add 03_visualize_study_cohorts.qmd
```
::::

Next, add a commit with an appropriate commit message. 

:::: {.shell}
```bash
git commit -m "Add new file '03_visualize_study_cohorts'"
```
::::

:::

Now that we've started tracking our work with Git, let's start by considering the first question: how do these studies compare in terms of the number of observations? 

Use the `dplyr` functions `group_by()` and `count()` to view the number of observations per study.

::: {.rcode}
```{r}
clean_metadata %>%
  group_by(Study) %>%
  count(name = "total")
```
:::

<details>
<summary>A note on the pipe operator `%>%`</summary>
The pipe operator `%>%` chains R commands together. It takes the preceding object and inputs it as the first argument to the function that follows it. It often is used to make code more readable and to avoid using intermediate objects or nested functions. The code above is equivalent to both of these expressions:

::: {.rcode}
```{r}
grouped_data <- group_by(clean_metadata, Study)
count(grouped_data, name = "total")
```
:::

::: {.rcode}
```{r}
count(group_by(clean_metadata, Study))
```
:::

The pipe operator that we use, `%>%`, is part of the `magrittr` R package. There is also a pipe operator in base R, `|>`. In most cases, these two pipe operators will work equivalently. 

::: {.rcode}
```{r}
clean_metadata |>
  group_by(Study) |>
  count(name = "total")
```
:::

</details>

We can see that there is quite the spread of study sizes. Let's also break this down by whether observations are from infants vs mothers. Whether the observation comes from an infant or a mother is contained in the variable `Category`.

**Your turn**: Modify the code above to count the number of observations by `Study` and by `Category`, and save the results as a data frame named `sample_counts`. **Add this code to your .qmd file.**

<details>
<summary>View code:</summary>
::: {.rcode}
```{r}
sample_counts <- clean_metadata %>%
  group_by(Study, Category) %>%
  count(name = "total")
sample_counts
```
:::
</details>

Now that we've looked at this data in a table format, let's make a plot so that we can compare this data visually. We'll make a bar plot, with separate bars for each study and for observation category (infant vs mother). **Include this code in your .qmd document** to make a bar plot.

::: {.rcode}
```{r}
sample_counts %>%
  ggplot(aes(x = Study, y = total, fill = Category)) +
  geom_col(position = position_dodge())
```
:::

While this plot looks good, it would be visually easier to compare if the studies were ordered by the number of observations. We can make a vector that orders the studies from the least observations to the most observations by using the `dplyr` functions `group_by()`, `summarise()`, `arrange()`, and `pull()`. 

<details>
<summary>More information about these `dplyr` functions:</summary>

| `dplyr` function | operation | 
| ---------------- | --------- |
| `group_by()` | groups rows by a variable | 
| `summarise()` | computes table of summaries | 
| `arrange()` | sorts rows by one or more variables | 
| `pull()` | pulls a variable from the data frame as a vector|

</details>

**Include this code in your .qmd document** to make a new vector for the desired study order.

::: {.rcode}
```{r}
study_order <- sample_counts %>%
  group_by(Study) %>%
  summarise(total_both = sum(total)) %>%
  arrange(total_both) %>%
  pull(Study)
study_order
```
:::

We can now order the bar plot by increasing sample size. **Include this code in your .qmd document** to make the bar plot with increasing bar heights.

::: {.rcode}
```{r}
sample_counts %>%
  mutate(Study = factor(Study, levels = study_order)) %>%
  ggplot(aes(x = Study, y = total, fill = Category)) +
  geom_col(position = position_dodge()) 
```
:::

Finally, we can update the colors of the plot and add labels. **Include this code in your .qmd document** to refine your bar plot.
 
::: {.rcode}
```{r}
sample_counts %>%
  mutate(Study = factor(Study, levels = study_order)) %>%
  ggplot(aes(x = Study, y = total, fill = Category)) +
  geom_col(position = position_dodge()) +
  theme_minimal() + 
  labs(y = "Number of samples",
       x = "Study",
       title = "Number of Samples by Study and Category") + 
  theme(plot.title = element_text(hjust = 0.5))
```
:::

We will now use `ggsave()` to save this plot in the figures folder. **Include this code in your .qmd document** to save this bar plot with the file name "study_sizes_by_category.png" within the "figures" folder.

::: {.rcode}
```{r, eval = FALSE}
ggsave("figures/study_sizes_by_category.png", height = 6, width = 9)
```
:::

This is another great time to commit your current work.

::: {.panel-tabset}

#### GitHub Desktop 

Navigate back to GitHub Desktop. You should now see that "03_visualize_study_cohorts.qmd" has a check in the checkbox (meaning it is staged) and has an orange box next to it, which means the file has been modified. Add an appropriate commit message and commit this file. 

#### Command line 

Even though Git is already tracking "03_visualize_study_cohorts.qmd", we still need to use `git add` to tell Git to stage this file for this commit. 

:::: {.shell}
```bash
git add 03_visualize_study_cohorts.qmd
```
::::

Next, add a commit with an appropriate commit message. 

:::: {.shell}
```bash
git commit -m "Add plot of study sizes to '03_visualize_study_cohorts'"
```
::::

:::

Next we will consider the second question: how do these studies compare in terms of when they took samples after the birth of the infants? You may have noticed in the script "02_wrangle_data.R" that two studies took samples from mothers during pregnancy. This information is contained in the variable `sampling_before_birth`. We would like to exclude these observations from our plots, because their `sampling_day` is counted from the beginning of pregnancy and not since birth. 

**Your turn**: create a data frame `observations_after_birth` that does not include these observations that were taken before birth. **Include this code in your .qmd document.**

<details>
<summary>Hint:</summary>
::: {.rcode}
The function `filter()` will remove rows that match a certain logical condition. 
:::
</details>

<details>
<summary>View code:</summary>
::: {.rcode}
```{r}
observations_after_birth <- clean_metadata %>%
  filter(!sampling_before_birth) 
```
:::
</details>

Now, we will make a histogram of sampling day, where we separate the data by `Category` and `Study`. There are multiple ways to do this, but we will color the bars by `Category` and then use `facet_wrap()` to make separate panels for each study. **Include this code in your .qmd document** to make histograms of sampling day for each study. 

::: {.rcode}
```{r}
observations_after_birth %>%
  ggplot(aes(x = sampling_day, fill = Category, group = Category)) +
  geom_histogram(position = position_dodge(), bins = 20) +
  facet_wrap(~Study, scale = "free_y") 
```
:::

Notice that because we set `scale = "free_y"` in the `facet_wrap()`, each of the y-axes of the panels is different. Let's order the studies in the same order that we used in the bar plot earlier. We will also update the theme and add axes labels. **Include this code in your .qmd document** to order the studies and refine your histograms. 

::: {.rcode}
```{r}
observations_after_birth %>%
  mutate(Study = factor(Study, levels = study_order)) %>%
  ggplot(aes(x = sampling_day, fill = Category, group = Category)) +
  geom_histogram(position = position_dodge(), bins = 20) +
  facet_wrap(~Study, scale = "free_y") +
  labs(x = "Days since birth at sampling",
       y = "Count",
       title = "Distribution of sampling day across studies") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
:::

<details>
<summary>More information about this warning message.</summary>
`ggplot` will give you warning messages when there are data in the dataframe that it cannot include in a plot due to missing information. In this warning, we can see that 7 observations in the data (representing samples taken) cannot be plotted because of non-finite values. Typically, this means there are observations with missing data. We can investigate what is missing with the following:

::: {.rcode}
```{r}
observations_after_birth %>%
  # only keep rows that have NA values for Study, sampling_day, or Category
  filter(is.na(Study) | is.na(sampling_day) | is.na(Category)) %>% 
  # view a smaller subset of variables
  select(Study, SampleID, PersonID, sampling_day, Category)
```
:::

Here we can see that there are 6 observations from the Backhed study and one observation from the Shao study for which `sampling_day` is missing. Therefore, these observations cannot be included in the plot.

</details>

Finally, we will save this plot in the figures folder. **Include this code in your .qmd document** to save this plot with the file name "sampling_timepoints.png" within the "figures" folder.

::: {.rcode}
```{r, eval = FALSE}
ggsave("figures/sampling_timepoints.png", height = 6, width = 9)
```
:::

Now that we have two plots, go back through your file "03_visualize_study_cohorts.qmd" to make sure that it contains all relevant code to make these two plots, and that there are explanations before each code chunk. Render your file to confirm that when rendered it will create these two plots. 



Since you've performed the requested analysis, make sure to document any new files in this repository in the README.md file. Specifically, describe what the file "03_visualize_study_cohorts.qmd" does, and describe the two plots that are now saved in the figures folder. After doing this, make another commit and push your work. 

::: {.panel-tabset}

#### GitHub Desktop 

Navigate back to GitHub Desktop. You should now see the modifications that you've made to "03_visualize_study_cohorts.qmd", as well as the added .png files for the figures. Add an appropriate commit message and commit. Then click "Push origin", to push these changes that you've made locally to your remote repository. 

#### Command line 

We need to now make sure to stage "03_visualize_study_cohorts.qmd" and the figures that we've added. We can use `git add .` to tell Git to track all files that have changes in this repository. 

:::: {.shell}

```bash
git add .
```

::::

Next, add a commit with an appropriate commit message. 

:::: {.shell}

```bash
git commit -m "Add plot of sampling days to 
'03_visualize_study_cohorts' and document new files in README"
```

::::

Finally, {{< glossary "push">}} these changes that you've made locally to your remote repository. Because this is your first time pushing from a new branch, you will need to tell Git where you want to push to, adding the flag `--set-upstream origin branch_name`. 

:::: {.shell}

```bash
git push --set-upstream origin study_comparison
```

::::

:::

***
