<!--
Describe the biological motivation for this case study.
Include relevant explanatory images and references throughout.
See this previous case study as an example: https://www.opencasestudies.org/ocs-bp-co2-emissions/#Motivation
Images or videos may be helpful so this includes an example of including an image.
-->

# **Motivation**
***

In this case study, we ask you to imagine that you are working with a collaborator on a scientific project. You receive the following email, with a request for your help performing a preliminary data analysis. Don't worry if you don't understand the context of this analysis right now, background will be given later in the case study. 

```{r, fig.alt="Motivation email asking for cohort comparison analysis", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3d9461be06f_0_0#slide=id.g3d9461be06f_0_0")
```

If you were to start working on this project with your collaborator, it may look something like this.

```{r, fig.alt="Back and forth over email without version control", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3e1f8f72664_0_368#slide=id.g3e1f8f72664_0_368")
```
```{r, fig.alt="Back and forth over email without version control", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3e1f8f72664_0_429#slide=id.g3e1f8f72664_0_429")
```


Between your own edits as you work on this project and the back-and-forth with your collaborator, it will quickly become challenging to keep track of all of the files and changes to them over time. A solution to this is {{< glossary "version control" >}}! 

```{r, fig.alt="Back and forth over email without version control", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3e1f8f72664_0_444#slide=id.g3e1f8f72664_0_444")
```

In this case study, you will learn how to use {{< glossary "Git" >}} and {{< glossary "GitHub" >}} to track this project and perform the requested comparison of studies used in the {{< glossary "meta-analysis" >}}. You will also learn how version control can facilitate collaboration and make a scientific project more {{< glossary "reproducible" >}}. 

You will learn more about version control, Git, and GitHub, in the [Introduction to Git and GitHub Section](https://www.opencasestudies.org/ocs-bio-version-control/#introduction-to-git-and-github) below. Much of the information in that section is described in the following papers. They both provide a helpful introduction to version control with Git and GitHub. 

:::{.reference_box}
::::{.reference_box_header}
Version control resources:
::::
::::{.inner_block}
- Blischak, J. D., Davenport, E. R., & Wilson, G (2016). [A Quick Introduction to Version Control with Git and GitHub](https://doi.org/10.1371/journal.pcbi.1004668), PLoS Computational Biology, 12(1).  
  - this paper provides an overview of Git and GitHub for scientists who primarily work with the command line
- Bryan, J (2018). [Excuse Me, Do You Have a Moment to Talk About Version Control?](https://doi.org/10.7287/peerj.preprints.3159v2) The American Statistician, 72(1), 20-27. 
  - this paper provides an overview of Git and GitHub for scientists and data scientists who primarily work with R and RStudio
::::
:::
  
***
