<!--
Describe the biological motivation for this case study.
Include relevant explanatory images and references throughout.
See this previous case study as an example: https://www.opencasestudies.org/ocs-bp-co2-emissions/#Motivation
Images or videos may be helpful so this includes an example of including an image.
-->

# **Motivation**
***

In this case study, we ask you to imagine that you are working with a collaborator on a scientific project. You receive the following email, with a request for your help performing a preliminary data analysis. Don't worry if you don't understand the context of this analysis right now, background will be given later in the case study. 

```{r, fig.alt="Motivation email asking for cohort comparison analysis", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3d9461be06f_0_0#slide=id.g3d9461be06f_0_0")
```

If you were to start working on this project with your collaborator, it may look something like this.

```{r, fig.alt="Back and forth over email without version control", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3e1f8f72664_0_368#slide=id.g3e1f8f72664_0_368")
```
```{r, fig.alt="Back and forth over email without version control", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3e1f8f72664_0_429#slide=id.g3e1f8f72664_0_429")
```


Between your own edits as you work on this project and the back-and-forth with your collaborator, it will quickly become challenging to keep track of all of the files and changes to them over time. A solution to this is version control! 

```{r, fig.alt="Back and forth over email without version control", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3e1f8f72664_0_444#slide=id.g3e1f8f72664_0_444")
```

In this case study, you will learn how to use Git and GitHub to track this project and perform the requested comparison of studies used in the meta-analysis. You will also learn how version control can facilitate collaboration and  make a scientific project more reproducible. 
  
***
