# **Continued Learning**
***

:::{.next_steps_box}
::::{.next_steps_box_header}
Common Next Steps
::::
::::{.inner_block}
Now that you've learned how to how to make and use version-controlled repositories, you are ready to use Git and GitHub for your work! Common next steps include:

- [adding collaborators to your repository](https://docs.github.com/en/repositories/managing-your-repositorys-settings-and-features/repository-access-and-collaboration/inviting-collaborators-to-a-personal-repository)
- making more branches for other components of your analysis
- using [GitHub issues](https://docs.github.com/en/issues/tracking-your-work-with-issues/learning-about-issues/about-issues) to keep track of to-do's related to your GitHub repository 
::::
:::

## Reinforcement Exercises

A good way to reinforce these version control skills is to make a current scientific project that you've been working on into a Git tracked repository, and connecting it to a remote GitHub repository. See if you can integrate the Git workflow that you've learned here into your scientific process. Even if you are the only one working on this project, this is a good way to practice skills. This will get you into the habit of staging, committing, and pushing changes as you work, and working on different branches for different components of a project. 

## Advancement Exercises

### Beginner

#### Moving work in progress into a new branch 

In this exercise, we will follow a process where we start to make changes to the main branch, but then realize that we would rather keep this work in a separate branch and only merge it into the main branch of the repository through a pull request after code review. 

Complete the following steps:

- Open repository for this case study
- Check out the main branch
- Add a new line to the README (add whatever you'd like here!)
- Follow the instructions on the [troubleshooting section](https://www.opencasestudies.org/ocs-bio-version-control/#troubleshooting-git-and-github) to move this update to the README to a new branch called updated_readme
- Commit and push your changes to a remote updated_readme branch
- Submit another pull request on GitHub to contribute this change into the main branch. 

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
What are some advantages to making changes to a repository within a new branch and using a pull request to merge these changes into the main branch? 

<details><summary>Answer</summary>

There are several advantages to this workflow: 

- Using a new branch lets you try new things without disrupting a stable version of the code and project
- Multiple people can work on different components of a project on their own branches, and then merge the branches into main when they are ready to contribute their work
- A branch and pull request lets you group several related commits together, and describe in more detail in the pull request why you made them 
- A pull request lets you request a code review from a collaborator before merging your work into main, making you less susceptible to mistakes in your code

</details>
::::
:::

### Intermediate

#### Going through a project's history 

In this exercise you will search through the history in a GitHub repository to find when a change took place. 
Open this [GitHub repository](https://github.com/opencasestudies/ocs-bio-version-control) which contains the contents of this case study. Feel free to spend some time poking around here, all of the content that is used to create different components of this case study are stored in this repository. 

Your goal is to find the pull request number and commit tag from when the learning objectives map table was filled in within the case study (the learning objectives table can be found [here](https://www.opencasestudies.org/ocs-bio-version-control/#objectives-map)). Do this by searching through the files, commits, or pull requests in this repository. 

<details>
<summary>Hints:</summary>

There are multiple ways to search for this change:

- Open the file _lo_map.qmd and click the "History" button to see a history of commits that affected this file.
- Go to the "Pull requests" tab, click on the "Closed" pull requests, and search a relevant search term. Once you find the correct pull request, look at the commits that make up that pull request.   

</details>

<details>
<summary>Answer:</summary>

Content was added to the learning objectives map in [pull request 18](https://github.com/opencasestudies/ocs-bio-version-control/pull/18), within the commit tagged [34a4550](https://github.com/opencasestudies/ocs-bio-version-control/commit/34a455071ec726470567932402f9496650ec46d6). 

</details>

### Advanced

#### Forking

So far, you've created repositories to work with. This means that GitHub considers you the repository owner. If you are working with collaborators who you want to be able to push to your GitHub repo, you can add them to your repository with write permissions (you can also make them a maintainer or admin to have additional abilities). However, let's consider a different situation. You want to contribute to a GitHub repository, but you do not have permission to push to this repository. This is where forks come in. 

:::{.definition_box}
::::{.definition_box_header}
Fork 
::::
::::{.inner_block}
A copy of a repository, often used to make changes when the user does not have write access to the original repository. 
::::
:::

We will call the original repository (which you do not have write permissions for) the upstream. Earlier, when you cloned a GitHub repository, you made a local version of that repository. In contrast, when you fork from the upstream, you make a remote copy of the GitHub repository, that you own with your GitHub account. You can then clone this fork, so that you can work locally. 

```{r, fig.alt="Fork workflow, making local repo", echo = FALSE, fig.align = "center"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3f139628ce6_0_369#slide=id.g3f139628ce6_0_369")
```

When you want to propose changes back to the upstream repository, you can submit a pull request from your fork into the upstream. 

```{r, fig.alt="Fork repo, integrating local changes to remote", echo = FALSE, fig.align = "center"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1TEi1D35oiiC21mYR0cKjWuOOpuxm3fTGyCXSqFmjL3I/edit?slide=id.g3f139628ce6_0_399#slide=id.g3f139628ce6_0_399")
```

##### Practice making and using a fork

To practice using a fork, go to [this GitHub repository](https://github.com/opencasestudies/version_controlled_infant_gut_analysis). You'll see that this repository contains the files that you originally downloaded when you started working on this infant gut analysis project. 

Look for the button that says "Fork" on the top right, above the "Code" button that you previously used to clone a repository. Click on "Fork". 

This will bring you to a page that is very similar to when you had created a new repository at the beginning of the case study. However, there are a few differences. You'll see now that you are the owner of this repository, and by default it has the same name as the repository that you are forking from. Click "Create fork". 

Once you create the fork, you'll be brought to this repository. Now, follow the same process that you used previously to clone this repository (*note: make sure that you are cloning the fork, and not the upstream repository!*). Open the repository locally. Make a copy of the file "03_visualize_study_cohorts.qmd", and add it to your local version of your forked repo. Follow the workflow to stage this file, commit it, and push it. 

After you push, you should now see the file on GitHub in your forked repo, along with a notification that this branch is a commit ahead of the upstream repository. There will be a button with the option to "Contribute" these changes. Click that button. This will take you to a page on the upstream repository where you can make a pull request from your fork. 

Fill in the title and description of your pull request. 

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}

How does this pull request from a fork to an upstream repository look different from the pull request that you made earlier, between two branches in the same repository? How does it look similar? 

::::
:::

Click "Create pull request" and inspect your pull request. 

Finally, please close this pull request, as we do not actually want to include these changes in this repository. 

***
