# **Troubleshooting Git and GitHub**
***

Most new and experienced users of {{< glossary "Git">}} and {{< glossary "GitHub">}} will run into problems occasionally. Luckily, {{< glossary "version control">}} is designed so that these problems can be solved, it just may take some time to get used to. There are many helpful [resources](https://dangitgit.com/en) online for common Git and GitHub problems, and AI chatbots can provide advice (please see our AI disclaimer below for advice on using AI for version control!). In this section, we will mention a few common Git and GitHub problems and provide solutions. 

<details> <summary> I {{< glossary "commit" display="committed">}} my work but I made a mistake in my most recent commit message that I would like to fix. </summary>

:::{.panel-tabset}
#### GitHub Desktop

Navigate to the repository in GitHub Desktop and go to the "History" pane. Right-click on the commit, and you should see the option "Amend Commit...", click on this option. 

If you have not yet pushed your work, you can then update your commit message and click "Commit", and then continue working.

However, if you have already pushed your work, then you will need to amend the commit message and then force push. This should only be done carefully, because it will rewrite the commit history of the branch for any other collaborator using that branch. Therefore, it is suggested to only do this for branches that you are working on alone, or to alert your collaborator before performing this operation and having them reset their branch after you force push. 

#### Command line

If you have committed but not yet {{< glossary "push" display="pushed">}} your work, you can open the command line and do the following:

:::: {.shell}
```bash
git commit --amend -m "Corrected commit message"
```
::::

If you have already pushed your work, you will need to run the previous line, and then push those changes.

:::: {.shell}
```bash
git push --force-with-lease
```
::::

The `--force-with-lease` option will push your change to the remote repository, unless the remote branch has new commits. If there are already new commits, this will fail, to avoid disrupting the version history of other users. If you would like to change the message anyway, you can replace `--force-with-lease` with `--force`. 

:::

</details>

<details> 
<summary>I've created a merge conflict. What is this? How do I fix it?</summary>

Typically, Git is able to automatically {{< glossary "merge">}} changes from one {{< glossary "branch">}} into another. However, a merge conflict happens when there are competing changes between the two branches, and Git needs you to manually decide how to resolve these differences. A merge conflict may arise in a scenario in which you made a branch from main, worked on that branch, and by the time you submit a pull request back to main, someone else has already committed to main and updated the same file as you have. 

:::{.definition_box}
::::{.definition_box_header}
Merge conflict
::::
::::{.inner_block}
When Git cannot automatically reconcile differences between branches it is attempting to combine, typically because the same part of a file (or closely related parts) were edited in incompatible ways, or because one branch modified something the other removed.
::::
:::

You will know that you have a merge conflict when Git attempts to combine changes (for example, during `git merge` or `git pull`) and reports that it cannot automatically resolve some differences. 

:::{.panel-tabset}
#### GitHub Desktop

You will see a pop-up window that says "Resolve conflicts before Merge", along with a list of files with conflicts. You will need to open the files in a text editor in order to manually resolve the merge conflicts. 

#### Command line

When you try to run `git pull`, you'll get a response that includes: "Automatic merge failed; fix conflicts and then commit the result". 

:::

Recall the hypothetical branch workflow from earlier in the case study.

::: {.panel-tabset}
#### Step 1

```{r, fig.alt="branch workflow: start with Main branch", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f9c55c112a_0_733#slide=id.g3f9c55c112a_0_733")
```

#### Step 2

```{r, fig.alt="branch workflow: add Plotting branch", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f9c55c112a_0_774#slide=id.g3f9c55c112a_0_774")
```

#### Step 3

```{r, fig.alt="branch workflow: Commit to Plotting branch", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f9c55c112a_0_796#slide=id.g3f9c55c112a_0_796")
```

#### Step 4

```{r, fig.alt="branch workflow: Separately, commit to main branch", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f9c55c112a_0_818#slide=id.g3f9c55c112a_0_818")
```

#### Step 5

```{r, fig.alt="branch workflow: merge Plotting back into Main", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f9c55c112a_0_840#slide=id.g3f9c55c112a_0_840")
```

:::

Here is an example of a file from the branching workflow that will result in a merge conflict. 

```{r, fig.alt="file before and after merge conflict", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1QjS4NwVgB_oZX5YCHKCzVe9x4XRWhNv-N7iP26JXHXo/edit?slide=id.g3f9c55c112a_0_874#slide=id.g3f9c55c112a_0_874")
```

When you open a file with a merge conflict, you will see that conflict markers have been added. For lines that are in conflict, you'll see "<<<<<<< HEAD" and "=======". The code in between these two markers is the code that is contained locally in the branch that you are working with. Lower in the file, you'll see ">>>>>>> incoming_change_tag", where "incoming_change_tag" references the commit or branch that identifies the incoming change. You need to manually edit the file to remove the tags, and to include whatever content you would like (either the content in your local branch, the content from the remote commit that is being pulled, or a combination of the two). Once you've edited this file, you should save and stage the file. 

In some cases (such as when one branch deletes a file and another edits it), you may not see conflict markers inside a file. Instead, you will need to either remove the file if you want it to be deleted, or stage the file if you would like it to be kept. 

After you resolve and stage all files that had merge conflicts, you should commit these changes. This will resolve your merge conflict. 

You can read more about when you might merge branches versus using a rebase operation to integrate changes in [this article](https://medium.com/@Spritan/merge-vs-rebase-vs-fast-forward-mastering-git-integration-strategies-3ba8e652d93e). 

</details>

<details> <summary> I've accidentally been working in the main branch, but I actually want my work to all go into a new branch. How do I switch branches and move this work with me?   </summary>

Luckily, this is an easy problem to resolve!

:::{.panel-tabset}
#### GitHub Desktop

If you have made changes locally but **haven't yet committed**, you can do the following: 

Open GitHub Desktop, and create a new branch. If you have changes in main that haven't been committed, GitHub Desktop will open a pop-up to ask you if you'd like to leave your changes on main or bring your changes to the new branch. Choose to bring the changes to the new branch. 

If you have made changes locally and **have committed**, you can follow the steps above and then additionally do the following: 

Switch back to the main branch, and click on History in the upper left panel. Then, right-click on the commit that you would like to return to in your main branch and choose "Reset to Commit..." from the right-click menu.

#### Command line

If you have made changes locally but haven't yet committed, you can do the following: 

Use the following command, replacing "new_branch_name" with your desired branch name. 

:::: {.shell}
```bash
git checkout -b new_branch_name 
```
::::

If you have made changes locally and have committed, you can follow the steps above to create a new branch, and then additionally do the following to revert main back to match the remote version of main: 

:::: {.shell}
```bash
git reset --hard origin/main
```
::::

However, make sure that you only `git reset --hard` after you've created the new branch and moved your commits, otherwise you risk losing your updated work.

Alternately, if you have only edited a few files, you could just run `git restore --hard` for those files. This reverts the files to their previous versions, instead of reverting the full branch. If you have edited files called "file1.txt" and "file2.txt" and want to revert them to the versions saved in main, you could do the following:

:::: {.shell}
```bash
git fetch origin
git restore --source=origin/main file1.txt file2.txt
```
::::


:::

</details>

<details> <summary> I've staged and committed a file, but I don't want to track it anymore. How can I remove it? </summary>

This is another common problem. We'll first tell Git to stop tracking the file (but keep it locally), and then we will add it to the .gitignore file to avoid tracking it with Git in the future. First, we will stop tracking the file. 

:::{.panel-tabset}
#### GitHub Desktop

Make sure you are starting in the correct branch that you have been working in. Click "Repository -> Open in Terminal". This will open a command line for you to work with. In this command line, use the following command, replacing "path_to_file" with the path to the file to stop tracking, including the file name. 

:::: {.shell}
```bash
git rm --cached "path_to_file" 
```
::::

Next, open the .gitignore file and add the file to the list of files to avoid tracking. 

Then, return to GitHub Desktop and commit these changes. 

#### Command line

Use the following command, replacing "path_to_file" with the path to the file to stop tracking, including the file name. 

:::: {.shell}
```bash
git rm --cached path/to/file
```
::::

Next, open the .gitignore file and add the file to the list of files to avoid tracking. 

Then, return to the command line and stage and stage and commit these changes. 

:::: {.shell}
```bash
git add .gitignore
git commit -m "removing file_name from Git tracking and adding it to .gitignore" 
```
::::

:::

</details>

<details> <summary> I've accidentally committed a large file and when I try to push I get an error from GitHub saying that the file is too big. How do I resolve this?  </summary>

This is a common error that can be a little tricky to resolve. You need to remove this file from Git tracking and also amend the commits that include the file. We will assume that the large file was added in the most recent commit. If it was added multiple commits previously, then it can still be removed but will include additional steps that we will not cover here (see [here](https://www.git-tower.com/learn/git/faq/git-filter-repo) for a workflow for removing a file from multiple commits). 

:::{.panel-tabset}
#### GitHub Desktop

Make sure you are starting in the correct branch that you have been working in. Click "Repository -> Open in Terminal". This will open a command line for you to work with. In this command line, use the following command, replacing "path_to_file" with the path to the file to stop tracking, including the file name. 

:::: {.shell}
```bash
git rm --cached "path_to_file" 
git commit --amend 
```
::::

Including the `git commit --amend` rewrites the most recent commit as if the large file was not added. Don't close the terminal yet, we will come back to it. 

Next, open the .gitignore file and add the file to the list of files to avoid tracking. In GitHub Desktop, stage the .gitignore, but do not commit it yet. 

Return to the terminal. Repeat the `git commit --amend` command, which will add this update to the .gitignore to the most recent commit. 

:::: {.shell}
```bash
git commit --amend 
```
::::

Then, return to GitHub Desktop and commit these changes. 

#### Command line

Use the following command, replacing "path_to_file" with the path to the file to stop tracking, including the file name. Including the `git commit --amend` rewrites the most recent commit as if the large file was not added.

:::: {.shell}
```bash
git rm --cached path/to/file
git commit --amend
```
::::

Next, open the .gitignore file and add the file to the list of files to avoid tracking. 

Then, return to the command line and stage and stage and commit these changes. Note that we include `git commit --amend` a second time to add the changes to the .gitignore file to the previous commit. 

:::: {.shell}
```bash
git add .gitignore
git commit --amend
git commit -m "removing file_name from Git tracking and adding it to .gitignore" 
```
::::

:::

The best way to avoid this problem is when you add large files to a folder that is tracked with Git, immediately add them to the .gitignore. If you really do want to track them with version control, consider using [Git Large File Storage](https://docs.github.com/en/repositories/working-with-files/managing-large-files/configuring-git-large-file-storage) (LFS). 

</details>

<details> <summary> I'm in the middle of working on one branch, but I need to pause and make a quick update on another branch. How do I move between branches if I'm not ready to commit my current work? </summary>

This is a common occurrence when you're working on a project with multiple branches. Imagine that you are deep in work for a feature that you're adding to an analysis, when you collaborator asks you to make a quick update to a difference branch that they've been working on. You aren't yet ready to commit the changes that you've been working on. This is a great time to use Git stash! 

:::{.panel-tabset}
#### GitHub Desktop

Open GitHub Desktop in the current branch that you're working with. In the top toolbar, click "Branch" -> "Stash All Changes". This will take all of the changes in this branch since your last commit and store them for you to come back to. Once you've done this, you should see no local changes to the branch. You can then move into the other branch you want to work on. 

Once you are ready to return to your original branch and continue your work, navigate back to that branch on GitHub Desktop. In the "Changes" panel, you should see the message "Stashed Changes". If you click on this, it will bring you to a page where you can click the "Restore" button to return your stashed changes to the branch. You can continue working from there. 

#### Command line

Open the command line to the branch that you've been working in. To save the changes in your current branch for you to later return to, use the following command:

:::: {.shell}
```bash
git stash
```
::::

After doing this, you can move to the other branch that you'd like to work on. Once you are ready to return to your original branch and continue your work, navigate back to that branch, and use the following command:

:::: {.shell}
```bash
git stash pop
```
::::

This command will restore your stashed changes to the current branch.

:::

To learn more about Git stash and additional options, you can read its [documentation from Git](https://git-scm.com/docs/git-stash).

</details>

You can learn more about other common Git mistakes and their fixes [in this article](https://www.creative-tim.com/blog/educational-tech/git-tutorial-fix-common-mistakes/).

:::{.ai_box}
::::{.ai_box_header}
AI use considerations
::::
::::{.inner_block}

<u>Context:</u>

When you ask an AI chatbot for assistance with version control, make sure to give appropriate background. Describe the specific task that you would like to do, or the problem that you are encountering. It can be helpful to include any error messages that you might be receiving from Git. For example, imagine that you've created a merge conflict. 

<u>Suggested Prompt:</u>

```{ .default .code-overflow-wrap }
I am working with a Git-controlled repository, in a branch called new_feature. I have made several changes locally, and have staged those files and committed them. I just went to pull changes from the remote version of this branch that my collaborator made recently, and I'm getting a merge conflict. Specifically, I've received the following message:

Auto-merging README.md
CONFLICT (content): Merge conflict in README.md
Auto-merging 01_import_data.R
CONFLICT (content): Merge conflict in 01_import_data.R
Automatic merge failed; fix conflicts and then commit the result.

How can I resolve this merge conflict?
```

<u>Validation:</u>
You should always be careful when taking advice on Git and GitHub from an AI chatbot. If you are asking for commands to use with Git for a situation that is very new to you, it could be helpful to do a quick search of the commands the chatbot tells you to use, to confirm that they do what you want to do. You could also ask multiple chatbots the same question. 

You should be especially careful when using agentic AI with Git and GitHub. Agentic AI systems can use external tools and execute tasks, often performing steps without human supervision. There are many git commands (and command line commands in general) that are challenging to reverse and can cause major problems to scientific projects. For example, running `git reset --hard` will reset a local branch to the remote version of the branch, removing all uncommitted changes, which could remove valuable content. Therefore, it is recommended to monitor the ways that agentic AI can make changes with version control and to prevent them from from performing potentially destructive actions without human supervision.
::::
:::

***
